---
title: "Appendix"
output: pdf_document
---

#Part1

##Loading dataset into dataframe
```{r}
library(readxl)
data <- read_excel(path = "/Users/sainivasrangaraju/Desktop/MMData.xlsx")
```

##New data frame with t and ln_y values
```{r}
lny_full_data <- data.frame(t = data$t,ln_y = log(data$y))
head(lny_full_data)

#Uncomment and run the below command to view completed data in lny_full_data
#lny_full_data
```

##Time selected for exponential peeling(t>4)
```{r}
time_selected <- which(data$t>4)
```

##Fitting a linear model
```{r}
lin1_model <- lm(ln_y~., lny_full_data, subset = time_selected)
lin1_model
```

##Extracting coefficients of the model and thus obtaining c1 and l1(lambda1)
```{r}
c1 <- exp(lin1_model$coefficients[1])
c1

l1 <- lin1_model$coefficients[2]
l1
```
##Finding y-c1*e^(-l1*t), the residual
```{r}
y_new <- data$y-c1*exp(l1*data$t)
head(y_new)

#Uncomment and run the below command to view completed data in y_new
#y_new
```

##Creating new dataframe with the log(residual_y)
```{r}
final_data <- data.frame(t = data$t,ln_y_res = log(y_new))
head(final_data)
```

##Fitting linear model to the residual log values
```{r}
lin2_model <- lm(ln_y_res~., final_data, subset = -time_selected)
lin2_model
```

##Extracting coefficients of the model and thus obtaining c2 and l2(lambda2)
```{r}
c2 <- exp(lin2_model$coefficients[1])
c2

l2 <- lin2_model$coefficients[2]
l2
```

##Creating new dataframe with final estimated values of y wrt t
```{r}
estimated_dataframe <- data.frame(y = c1*exp(l1*data$t)+c2*exp(l2*data$t),
                                  t = data$t)
head(estimated_dataframe)

#Uncomment and run the below command to view the  data in estimated_dataframe
#estimated_dataframe
```

##Graph for estimated vs actual y values
```{r}
library(ggplot2)
ggplot()+
  geom_point(data = data,aes(t,y, shape = "Actual y"))+
  geom_point(data = estimated_dataframe, aes(t,y, shape = "Estimated y"))+
  geom_line(data = estimated_dataframe, aes(t,y))+
  ggtitle("Estimated y vs Actual y")+
  labs(shape = "Legend Title")+
  scale_shape_manual(values = c("Actual y" = 17, "Estimated y" = 20))
```

##Sum of Squared Errors for the estimated vs actual
```{r}
SR <- (estimated_dataframe$y-data$y)^2
SSR <- sum(SR)
cat("Sum of Squared Errors for the estimated vs actual is ",SSR)
```

##Using Optimization function optim() to find the c1, l1, c2 and l2
```{r}
objective_function <- function(x) {
  SE = (data$y-x[1]*exp(x[2]*data$t)-x[3]*exp(x[4]*data$t))^2
  SSE = sum(SE)
  return(SSE)
}

# Initial guess
initial_guess <- c(0,0,0,0)

# Call optim
result <- optim(par = initial_guess, fn = objective_function, method = "BFGS")

# Display results
cat("\nMinimum found at:")
cat("\n\tc1 =", result$par[1], "\n\tl1 =", result$par[2],
    "\n\tc2 =", result$par[3], "\n\tl2 =", result$par[4])
cat("\nThe minimum SSE obtained is =", result$value, "\n")
```
##Creating dataframe for y values estimated by optimized paramaters
```{r}
optimized_dataframe <- data.frame(t = data$t,y = result$par[1]*exp(result$par[2]*data$t)+result$par[3]*exp(result$par[4]*data$t))
head(optimized_dataframe)
```

##Sum of Squared Errors for the actual vs y obtained by optimized parameters
```{r}
SE <- (optimized_dataframe$y-data$y)^2
SSE <- sum(SE)
cat("Sum of Squared Errors for the optimized vs actual is ",SSE)
```


##Graph for actual y vs estimated y vs y obtained by optimized parameters
```{r}
ggplot()+
  geom_point(data = data,aes(t,y, shape = "Actual y"))+
  geom_point(data = estimated_dataframe, aes(t,y, shape = "Estimated y"))+
  geom_point(data = optimized_dataframe, aes(t,y, shape="Optimized y"))+
  scale_shape_manual(values = c("Actual y" = 17, 
                          "Estimated y" = 20,"Optimized y" = 15))+
  ggtitle("Original y vs Exponential peeling's y vs Optimized parameters' y")
```

##Comparing the exponential peeling result vs optimized parameters result
The optimization function gave us the parameters values as: c1 = 5.132154,
l1 = -0.1046885, c2 = 4.941015 and l2 = -1.080367. 
These values are almost similar to the values obtained by exponential 
peeling, the values are c1 = 5.328934, l1 = -0.1107134, 
c2 = 4.447866 and l2 = -1.12763

It is also clear from the graph that the y-values(points) obtained by using
exponential peeling(points in purple) and by using optimized parameters
(points in cyan) overlap almost perfectly with the original y values(points in
orange)


#Part 2

##Loading dataset into dataframe
```{r}
data <- readxl::read_excel("/Users/sainivasrangaraju/Desktop/MM Part2.xlsx")
head(data)

#Uncomment the below code to view full data
#data
```

##Time selected for exponential peeling(t>50)
```{r}
time_split <- which(data$time > 50)
```

##New data frame with t, ln(x1) and ln(x2) values
```{r}
log_data <- data.frame(t = data$time, ln_x1 = log(data$x1), ln_x2 = log(data$x2))
log_data
```

##Fitting linear model to ln(x1) and t
```{r}
lin1_model <- lm(ln_x1~t, log_data, subset = time_split)
lin1_model
```

##Fitting linear model to ln(x2) and t
```{r}
lin2_model <- lm(ln_x2~t, log_data, subset = time_split)
lin2_model
```

##Obtaining lambda1, alpha*v11 and alpha*v21
```{r}
lambda1 <- (lin1_model$coefficients[2]+lin2_model$coefficients[2])/2
lambda1
```

```{r}
c1 <- exp(lin1_model$coefficients[1])
#c1 is alpha*v11
c1
```

```{r}
c2 <- exp(lin2_model$coefficients[1])
#c2 is alpha*v21
c2
```

##Plotting the linear models with log(x1) and log(x2)
```{r}
library(ggplot2)
ggplot(log_data, aes(x = t))+
  geom_point(aes(y = ln_x1, shape = "x1"))+
  geom_point(aes(y = ln_x2, shape = "x2"))+
  geom_abline(intercept = lin1_model$coefficients[1], slope = lambda1)+
  geom_abline(intercept = lin2_model$coefficients[1], slope = lambda1)+
  scale_shape_manual(values = c("x1" = 17, "x2" = 20))+
  labs(title = "x1 and x2 vs t",y="ln(x1) and ln(x2)", x = "t")
```

##Calculating the residuals of x1 and x2
```{r}
new_x1 <- data$x1 - c1*exp(lambda1*data$time)
new_x2 <- data$x2 - c2*exp(lambda1*data$time)
```

##Creating dataframe for log of the residuals of x1 and x2
```{r}
new_log_data<-data.frame(t = data$time,ln_x1 = log(new_x1),ln_x2 = log(new_x2))
new_log_data
```

##Fitting linear model to the ln(x1*)
```{r}
lin3_model <- lm(ln_x1~t, new_log_data, subset = -time_split)
lin3_model
```

##Obtaining lamda2
```{r}
lambda2 <- as.numeric(lin3_model$coefficients[2])
lambda2
```

##Obtaining beta*v12
```{r}
c3 <- as.numeric(exp(lin3_model$coefficients[1]))
c3
```

##Taking beta*v22 = -(beta*v12)(from Initial conditions)
```{r}
#Here we can't fit the linear model to the log of x2* as the x2* values are
#negative and the log will be not defined. To avoid this we skipped fitting
#linear model to the log(x2*) and calculated the value of beta*v22 from the
#initial condition i.e; the first entry in the data, which gives us: 
#beta*v22 = -(beta*v12)
c4 <- -c3
c4
```

##Calculating estimated values
```{r}
#time[-1] excludes the first entry i.e; the initial condition
estimated_x1 <- c1*exp(lambda1*data$time[-1])+c3*exp(lambda2*data$time[-1])
estimated_x2 <- c2*exp(lambda1*data$time[-1])+c4*exp(lambda2*data$time[-1])
```

##Creating dataframe with estimated values
```{r}
estimated<-data.frame(t=data$time[-1],x1=estimated_x1,x2=estimated_x2)
```

##Calculating Sum of Squared errors for both x1 and x2
```{r}
x1_sse <- sum((data$x1[-1]-estimated$x1)^2)
x1_sse
```
```{r}
x2_sse <- sum((data$x2[-1]-estimated$x2)^2)
x2_sse
```

##Plotting the estimated vs actual values of x1 and x2
```{r}
ggplot()+
geom_point(data,mapping=aes(x =time,y = x1, shape="Actual x1"))+
geom_point(estimated,mapping=aes(x =t,y = x1, shape="Estimated x1"))+
geom_point(data,mapping=aes(x =time,y = x2, shape="Actual x2"))+
geom_point(estimated,mapping=aes(x =t,y = x2, shape="Estimated x2"))+
geom_line(data,mapping=aes(x =time,y = x1))+
geom_line(estimated,mapping=aes(x =t,y = x1))+
geom_line(data,mapping=aes(x =time,y = x2))+
geom_line(estimated,mapping=aes(x =t,y = x2))+
scale_shape_manual(values = c("Actual x1" = 15, "Estimated x1" = 17,
                                "Actual x2" = 19, "Estimated x2" = 12))+
ggtitle("Estimated concentrations vs Actual concentrations")+
labs(x = "time", y = "Concentration(x)")
```
